Combinatorial Identities from Lagrange's Interpolation Polynomial

### 2021-01-07 spectral radius of matrix new method/thesis/version_12_28_2020.tex
010721 read

{ \Huge 有向圖譜半徑之簡易估算方法}~\\~\\

@@

{ \Huge A simple method on estimating spectral radius for some directed graphs}~\\~\\
~\\~\\~\\~\\~\\
\bigskip \bigskip\bigskip\bigskip\bigskip\bigskip
\bigskip \bigskip\bigskip\bigskip\bigskip\bigskip
\bigskip\bigskip\bigskip

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\renewcommand{\baselinestretch}{2} %行距
\thispagestyle{empty}
\begin{center}
{
\Large
有向圖譜半徑之簡易估算方法\\
A simple method on estimating spectral radius for some directed graphs\\~\\
\begin{tabular}{lccr}
Student: Ko-Han Chen  &&~~~& Advisor: Chih-Wen Weng\\
研究生:陳科翰　 &&~~~& 指導教授：翁志文~教授
\end{tabular}
}~\\

@@

%中文摘要
\pagenumbering{roman}
\begin{center}
{  \LARGE
有向圖譜半徑之簡易估算方法
\bigskip\bigskip\bigskip

@@

 矩陣的譜半徑為其特徵值絕對值的最大值，而一有向圖的譜半徑則定義為其鄰接矩陣之
 譜半徑。本論文給出一個尋找方陣譜半徑上界的方法，將此方法應用於有向圖的鄰接矩
 陣，我們可以簡易估算出此圖的譜半徑。
 \\
\bigskip

@@

%英文摘要
\begin{center}{\LARGE
A simple method on estimating spectral radius for some directed graphs
\bigskip\bigskip\bigskip}

@@

%\begin{abstract}
\addcontentsline{toc}{section}{Abstract (in English)}
The spectral radius of a square matrix is the largest magnitude of its eigenvalues. And the
 spectral radius of a directed graph is defined as the spectral radius of the corresponding
 adjacency matrix. In this paper, we give an approach to obtain an upper bound for the
 spectral radius of a square matrix. By applying this method on the adjacency matrix of a
 directed graph, we can estimate the spectral radius of this directed graph simply.

@@

Let $\mathbb{R}$ and $\mathbb{C}$ denote the
 field of real numbers and complex numbers, respectively.
 Let $C$ be an $n\times n$ real square matrix. If there is
 a nonzero column vector $u\in\mathbb{R}^n$ such that
 $Cu=\lambda u$ for some scalar $\lambda\in\mathbb{C}$,
 then the scalar $\lambda$ is called the eigenvalue of
 $C$ with corresponding eigenvector $u$. And the spectral
 radius of a matrix $C$ is the largest magnitude (or complex
 modulus) of its eigenvalues, denoted by $\rho(C)$. We are
 interested in the spectral radius of the following matrix
 associated with a simple directed graph.

\begin{defn}
Given an directed graph $G$, the $\textit{adjacency matrix}$ of
 $G$ is the square
 matrix $A = (a_{ij})$ indexed by vertices of $G$, and
     \[a_{ij} =\begin{cases}
        1, \text{if $i$ is adjacent to $j$}, \\
        0, \text{otherwise.}
            \end{cases}
     \]
\end{defn}

@@

@@

The following is Perron–Frobenius theorem, which provides a feature of
 nonnegative eigenvector to nonnegative matrices.

\begin{thm} \cite{prn_fros2} \label{thm:Perron_Frobenius}
    If $C$ is a nonnegative square matrix, then the spectral radius $\rho(C)$ is an
    eigenvalue of $C$ with a corresponding nonnegative right eigenvector and a
    corresponding nonnegative left eigenvector.
\end{thm}

We introduce a notation of submatrix, which is taken from some columns and some rows of a
 matrix.

\begin{defn}
    For a matrix $C=(c_{ij})$ and subsets $\alpha$, $\beta$ of row indices and column
    indices of $C$, respectively, we use $C[\alpha|\beta]$ to denote the
    submatrix of $C$ with size $ |\alpha| \times |\beta| $ that has entries $c_{ij}$ for $i\in \alpha$
    and $j\in\beta$.
\end{defn}

Our theory is based on the following theorem, which is from \cite{chang}, and we introduce two square matrices
 $P$ and $Q$ in the following theorem first. The matrix $P$ is a permutation
 matrix which is multiplied to the left side, and the matrix $Q$ is the sum of an elementary matrix and a
 certain binary matrix. In which $P$ generalize row permutation on cases of $C$ matrix, and $Q$ is the transform
 from $C'$ to $C'$, which is the first $n-1$ columns and sum of certain columns. We aim to find $C'$ such that
 $C'$ majors $C$, i.e. $C\leq C'$.

\begin{thm}\label{pre_thm}
    Let $C=(c_{ij})$, $C'=(c'_{ij})$, $P$ and $Q$ be $n\times n$ matrices.
Assume that
\begin{enumerate}[label=(\roman*)]
    \item \label{pre_thm_em1}  $PCQ\leq PC'Q$;
    \item \label{pre_thm_em2} there exist a nonnegative column vector $u=(u_1, u_2, \ldots, u_n)^T$  and a
    scalar $\lambda'\in \mathbb{R}$ such that $\lambda'$ is an eigenvalue of $C'$ with
    associated eigenvector $Qu$;
    \item \label{pre_thm_em3}  there exist a nonnegative row vector $v^T=(v_1, v_2, \ldots, v_n)$  and a scalar
    $\lambda\in \mathbb{R}$such that $\lambda$ is an eigenvalue of $C$ with associated  left
    eigenvector $v^TP$; and
    \item \label{pre_thm_em4} $v^TPQu>0.$
\end{enumerate}
    Then $\lambda\leq \lambda'$. Moreover, $\lambda=\lambda'$ if and only if
    \begin{equation}\label{pre0}
        (PC'Q)_{ij}=(PCQ)_{ij}\qquad \hbox{for~}1\leq i, j\leq n \hbox{~with~} v_i\ne 0 \hbox{~and~} u_j\ne 0.
    \end{equation}
\end{thm}

\begin{proof}
    Multiplying the nonnegative vector $u$ in theorem~\ref{pre_thm}
     assumption~\ref{pre_thm_em1},
     where $Qu$ is eigenvector of $C'$, 
     to the right of both terms of~\ref{pre_thm_em1},
    \begin{equation}\label{pre1}
       PCQu\leq PC'Qu=\lambda'PQu.
    \end{equation}

    Multiplying the nonnegative left eigenvector $v^T$ of $C$ for $\lambda$ in assumption
     ~\ref{pre_thm_em3} to the left of all terms  in (\ref{pre1}),
      where $v^TP$ is left eigenvector of $C$ for $\lambda$, thus we have
    \begin{equation}\label{pre2}
        \lambda v^TPQu=v^TPCQu\leq v^TPC'Qu=\lambda' v^TPQu.
    \end{equation}
        Now delete the positive term $v^TPQu$ by assumption \ref{pre_thm_em4} to obtain
        $\lambda\leq \lambda'$ and finish the proof of the first part.
        Assume that $\lambda=\lambda'$, so the inequality in (\ref{pre2}) is an equality.
        Especially $(PCQu)_i=(PC'Qu)_i$ for any $i$ with $v_i\not=0.$ Hence,
        $(PCQ)_{ij}=(PC'Q)_{ij}$ for any $i$ with $v_i\not=0$ and any $j$ with
        $u_j\not=0.$ Conversely, (\ref{pre0}) implies $$v^TPCQu=\sum_{i,j} v_i(PCQ)_{ij}u_j=
         \sum_{i,j} v_i(PC'Q)_{ij}u_j=v^TPC'Qu,$$ so $\lambda=\lambda'$ by (\ref{pre2}).
\end{proof}

@@

We use $[n-1]$ as notation of the set of elements from one to $n-1$, which is $\{1,2,...,n-1\}$.
Throughout fix $k\in [n-1]$, let $E_{kn}$ denote the $n\times n$ binary matrix with a unique $1$
appearing in the position $k,n$ of $E_{kn}$. Now we apply the previous Theorem \ref{pre_thm} with $P=I$ and

\begin{equation} \label{Q_1}
Q=I+E_{kn}=\begin{pmatrix}
1 &   &         &   & 0 \\
  & 1 &         &   &   \\
  &   & \ddots  &   & 1 \\
  &   &         & 1 &  \\
0 &   &         &     & 1
\end{pmatrix},
\end{equation}
so the matrix $PC'Q$ in assumption~\ref{pre_thm_em1} of Theorem~\ref{pre_thm} is
\begin{equation}\label{PC'Q}
  PC'Q=\begin{pmatrix}
         c'_{11} & c'_{12} & \cdots &  c'_{1k}+c'_{1n} \\
         c'_{21} & c'_{22} & \cdots &  c'_{2k}+c'_{2n} \\
         \vdots & \vdots & \ddots &  \vdots \\
         c'_{n1} & c'_{n2} & \cdots &  c'_{nk}+c'_{nn}
       \end{pmatrix},
\end{equation}
where $c'_{ij}$ denotes the $(i,j)$-entry of $C'$.

\begin{defn}%[$K$-rooted vector]
A column vector $v'=(v'_1,v'_2,\ldots,v'_n)^T$ is called {\it $K$-rooted}  if $v'_{j} \geq 0$ for $1 \leq  j \leq n$ and $v'_k\geq v'_n.$
\end{defn}

The following Lemma is immediate from the above definition.%[vector rooted lemma]

\begin{lem}\label{lem:rt_vec}
If $u=(u_1, u_2, \ldots, u_n)^T$ and $v'=(v'_1, v'_2, \ldots, v'_n):=Qu=(u_1,\ldots, u_{k-1},u_k+u_n, u_{k+1}, \ldots,  u_n)^T$, then
\begin{enumerate}[label=(\roman*)]
\item \label{lem:rt_vec:en1}$v'$ is $k$-rooted  if and only if $u$ is nonnegative;
\item $u_k>0$ if and only if $v'_k>v'_n$.
\end{enumerate}
% \qed
\end{lem}

Below is our first result, in which the first condition implies the first $n-1$ columns of $C$ major to
 the columns of $C'$, and the $(k,n)$-sum column of $C$ is also major to $C'$. The second and the third condition
 suggest that $C$ and $C'$ have nonnegative eigenvectors which are $K$-rooted. And the forth condition is simpler
  but with the same meaning with Theorem~\ref{pre_thm}
\begin{thm}\label{thm_main}
    Let $C=(c_{ij})$, $C'=(c'_{ij})$ be  $n\times n$ matrices.
Assume that
\begin{enumerate}[label=(\roman*)]
\item \label{thm_main:condition_i} $C[[n]|[n-1]]\leq C'[[n]|[n-1]]$ and $c_{ik}+c_{in}\leq c'_{ik}+c'_{in}$ for all $1\leq i\leq n$;
\item \label{thm_main:condition_ii} there exists a $K$-rooted vector $v'=(v'_1, v'_2, \ldots, v'_n)^T$ and a scalar $\lambda'\in \mathbb{R}$
such that $\lambda'$ is an eigenvalue of $C'$ with associated eigenvector $v'$;
\item \label{thm_main:condition_iii}there exists a nonnegative vector $v^T=(v_1, v_2, \ldots, v_n)$ and a scalar $\lambda\in \mathbb{R}$ such that $\lambda$ is an eigenvalue of $C$ with associated left eigenvector $v^T$;
\item \label{thm_main:condition_iv}$v^Tv'>0.$
\end{enumerate}
 Then $\lambda\leq \lambda'$.
Moreover, $\lambda=\lambda'$
if and only if
\begin{enumerate}[label=(\alph*)]
    \item \label{thm_main:equ_cond_a} $c_{ik}+c_{in}=c'_{ik}+c'_{in} \qquad$  for $1\leq i\leq n$ with $v_i\not=0$ and $v'_n\not=0;$
    \item \label{thm_main:equ_cond_b} $c'_{ij}=c_{ij}\qquad $for $1\leq i\leq n,~1\leq j\leq n-1, j \neq k $with $v_i\ne 0 $;
    \item \label{thm_main:equ_cond_c} $c'_{ik}=c_{ik} \qquad $  for $1\leq i \leq n$ and $ v'_{k}>v'_n$
\end{enumerate} %\qed
\end{thm}
%

@@

We are interested in the matrices $C'$ that have $K$-rooted eigenvectors.
Motivated by the condition (i) of Theorem 2.3, we provide the following two definitions.
This is the definition of $(k,n)$-sum.
\begin{defn}
    For an $n \times n$ matrix $C'=(c'_{ij})$, the $(k, n)$-sum vector of $C'$ is the vector
     of the sum of the $k$-th and  $n$-th columns of $C'$, where $k\leq n-1$.
\end{defn}

Note that the last column of $C'Q$ is the $(k, n)$-sum vector of $C'$.
Below is the definition of $K$-rooted matrix.
\begin{defn}\label{m_rooted}
    A  matrix $C'=(c'_{ij})$ is called {\it $K$-rooted} if its
     first $n-1$ columns and its $(k,n)$-sum vector are all $K$-rooted.
\end{defn}

\begin{rem}

    $$Q^{-1}=I-E_{kn}=\begin{pmatrix}
    1 &  & &   & 0 \\
    & 1 &        &  &  \\
    &  & \ddots  &  & -1 \\
    &  & &  1 &  \\
    0 &  &  &  & 1
    \end{pmatrix}.$$

\end{rem}

The following lemma shows that a $K$-rooted matrix has a $K$-rooted eigenvector.
\begin{lem}\label{lma_m_rooted}
    Let $C'=(c'_{ij})$ be an $n\times n$ nonnegative matrix. Then the following (i)-(iii) hold.
        \begin{enumerate}[label=(\roman*)]
            \item \label{lma_m_rooted_cond1} $C'$ is a $K$-rooted matrix if and only if $Q^{-1}C'Q$ is nonnegative,
            \item \label{lma_m_rooted_cond2}Assume that $C'$ is $K$-rooted and let $u$ be the nonnegative eigenvector of $Q^{-1}C'Q$
                corresponding to $\rho(C')$. Then $C'$ has a $K$-rooted eigenvector $v'=Qu$ for $\rho(C')$,
            \item \label{lma_m_rooted_cond3}$\rho(C')$ = $\rho(Q^{-1}C'Q)$.
        \end{enumerate}
\end{lem}

\begin{comment}
\begin{rem}
        (i) is immediate from Definition~\ref{m_rooted} and the observation that
        $$Q^{-1}=I-E_{kn}=\begin{pmatrix}
        1 &  & & &  & 0 \\
         & 1 &  &      &  &  \\
         &  & \ddots & &  & -1 \\
         &  &        & &  &  \\
          &  & & & 1 &  \\
        0 &  & & &  & 1 \\
        \end{pmatrix},$$
\end{rem}
\end{comment}
\begin{proof}
\begin{enumerate}
  \item[(i)] The matrix $Q^{-1}C'Q$ is
        $$\begin{pmatrix}
            c'_{11}     & c'_{12} & \cdots     & c'_{1\ n-1} & c'_{1k}+c'_{1n} \\
            \vdots      & \vdots  & \vdots     & \vdots      & \vdots\\
            c'_{(k-1) 1}   & c'_{(k-1)  2}           & \cdots     & c'_{(k-1) (n-1)} & c'_{(k-1) k}+c'_{(k-1) n} \\
            c'_{k1}-c'_{n1} & c'_{k2}-c'_{n2} &\cdots      &c'_{k (n-1)}-c'_{n (k-1)}& c'_{kk}+c'_{kn}-c'_{nk}-x'_{nn}\\
            c'_{(k+1) 1}   & c'_{(k+1) 2}           & \cdots     & c'_{(k+1) (n-1)} & c'_{(k+1) k}+c'_{(k+1)
             n} \\
            \vdots              & \vdots & \ddots              & \vdots & \vdots \\
            c'_{n1}     & c'_{n2} & \cdots             & c'_{n (n-1)} & c'_{nk}+c'_{nn} \\
        \end{pmatrix},$$

        which is nonnegative if and only if $C'$ is $K$-rooted by the definition of nonnegative matrix and $K$-rooted matrix.
  \item[(ii)] By Lemma~\ref{lem:rt_vec}, $v'=Qu$ is $K$-rooted. Since $Q^{-1}C'Qu=\rho(C')u$ by the assumption, we have
      $$Q^{-1} C' Q u  = Q^{-1} \rho(C') Qu  =\rho(C')u,$$
      which implies
        $$C'Qu=\rho(C')Qu.$$
  \item[(iii)] Since $C'$ and $Q^{-1}C'Q$ have the same set of eigenvalues, clearly
  $$\rho(C') = \rho(Q^{-1}C'Q).$$
\end{enumerate}
\end{proof}

\begin{lem}\label{l_diag}   %end of label
If a square matrix $C'$ has a $K$-rooted eigenvector corresponding to $\lambda'$, then $C'+dI$ also has
the same $K$-rooted eigenvector corresponding to $\lambda'+d,$ where $d$ is a constant and $I$ is the identity matrix with the same size of $C'$.
\end{lem}

\begin{thm}\label{thm:conclusion}
    Let $C$ be an $n\times n$ nonnegative matrix. For $1\leq i \leq n$ and $1\leq j\leq n-1$, choose $c'_{ij}$
    such that $c'_{ij}\geq c_{ij}$ and $c'_{kj}\geq c'_{nj}>0$, and choose $r'_i$ such that $r'_i\geq c_{ik}+c_{in}$, and
    $r'_k \geq r'_n$. Moreover choose $c'_{in}:=r'_i-c'_{ik}$. Then $\rho(C)\leq \rho(C')$, when $C'=(c'_{ij})$.
\end{thm}

>>>>

{ \Huge 有向圖譜半徑之簡易比較方法}~\\~\\

@@

{ \Huge A simple method on comparison between spectral radii of two directed graphs}~\\~\\
~\\~\\~\\~\\~\\
\bigskip \bigskip\bigskip\bigskip\bigskip\bigskip
\bigskip \bigskip\bigskip\bigskip\bigskip\bigskip
\bigskip\bigskip\bigskip

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\renewcommand{\baselinestretch}{2} %行距
\thispagestyle{empty}
\begin{center}
{
\Large
有向圖譜半徑之簡易比較方法\\
A simple method on comparison between spectral radii of two directed graphs\\~\\
\begin{tabular}{lccr}
Student: Ko-Han Chen  &&~~~& Advisor: Chih-Wen Weng\\
研究生:陳科翰　 &&~~~& 指導教授：翁志文~教授
\end{tabular}
}~\\

@@

%中文摘要
\pagenumbering{roman}
\begin{center}
{  \LARGE
有向圖譜半徑之簡易比較方法
\bigskip\bigskip\bigskip

@@

 矩陣的譜半徑為其特徵值絕對值的最大值，而一有向圖的譜半徑則定義為其鄰接矩陣之
 譜半徑。本論文給出一個比較方陣譜半徑的方法，將此方法應用於有向圖的鄰接矩
 陣， 我們可以簡易比較有向圖的譜半徑。
 \\
\bigskip

@@

%英文摘要
\begin{center}{\LARGE
A simple method on comparison between spectral radii of two directed graphs
\bigskip\bigskip\bigskip}

@@

%\begin{abstract}
\addcontentsline{toc}{section}{Abstract (in English)}
The spectral radius of a square matrix is the largest magnitude of its eigenvalues. And the
 spectral radius of a directed graph is defined as the spectral radius of the corresponding
 adjacency matrix. In this paper, we give an approach to compare the spectral radii of two
 nonnegative matrices. By applying this method on the adjacency matrix of a
 directed graph, we can compare the spectral radii of two directed graphs simply.

@@

Let $\mathbb{R}$ and $\mathbb{C}$ denote the
 field of real numbers and complex numbers, respectively.
 Let $C$ be an $n\times n$ real square matrix. If there is
 a nonzero column vector $u\in\mathbb{C}^n$ such that
 $Cu=\lambda u$ for some scalar $\lambda\in\mathbb{C}$,
 then the scalar $\lambda$ is called the eigenvalue of
 $C$ with corresponding eigenvector $u$. And the spectral
 radius of a matrix $C$ is the largest magnitude (or complex
 modulus) of its eigenvalues, denoted by $\rho(C)$. We are
 interested in the spectral radius of the following matrix
associated with a simple directed graph.

\begin{defn}
    Given a directed graph $G$, the $\textit{adjacency matrix}$ of $G$ is the square
    matrix $A = (a_{ij})$ indexed by vertices of $G$, and
     \[a_{ij} =\begin{cases}
        1, \text{if $ji$ is an arc in $G$}, \\
        0, \text{otherwise.}
            \end{cases}
     \]
\end{defn}

@@

Conversely, we can also define a directed graph $G$ from a given
 nonnegative $n\times n$ matrix $C=(c_{ij})$  by setting the vertex
 set $\{1, 2, \ldots, n\}$ and edge set $\{ij~\colon~c_{ij}>0\}$.
 The matrix $C$ is {\it irreducible} if the defined graph $G$ from
 $C$ is strongly connected.

@@

The following is Perron–Frobenius theorem, which provides a feature of
 nonnegative eigenvectors to nonnegative matrices.

\begin{thm} \cite{prn_fros2} \label{thm:Perron_Frobenius}
    If $C$ is a nonnegative square matrix, then the spectral radius $\rho(C)$ is an
    eigenvalue of $C$ with a corresponding nonnegative right eigenvector and a
    corresponding nonnegative left eigenvector.
    Moreover if $C$ is irreducible the above eigenvectors can be chosen to be positive. 
\end{thm}

 A well-known application of Theorem~\ref{thm:Perron_Frobenius}
 show that if matrix  $C'$ {\it majors} $C$ (in notation $C\leq C'$),
 i.e. $c_{ij}\leq c'_{ij}$ for all $ij,j$, then $\rho(C)\leq \rho(C')$.
Our main result shows that the assumption $C\leq C'$ can be a little loosen. 
Our theory is based on the following theorem, which is from \cite{chang}.

\begin{thm}\label{pre_thm}
    Let $C=(c_{ij})$, $C'=(c'_{ij})$  be $n\times n$ real matrices
     with real eigenvalues $\lambda$, $\lambda'$ respectively such
     that there exist $n\times n$ matrices $P$ and $Q$   satisfying the following (i)-(iv).
\begin{enumerate}[label=(\roman*)]
    \item \label{pre_thm_em1}  $PCQ\leq PC'Q$;
    \item \label{pre_thm_em2} an eigenvector $Qu$ of $C'$  associated with
     eigenvalue $\lambda'$ exists for some nonnegative column vector
      $u=(u_1, u_2, \ldots, u_n)^T$.
    \item \label{pre_thm_em3} a left eigenvector $v^TP$ of $C'$ associated
     with eigenvalue $\lambda$ exists for some nonnegative row vector
     $v^T=(v_1, v_2, \ldots, v_n)$; and
    \item \label{pre_thm_em4} $v^TPQu>0.$
\end{enumerate}
    Then $\lambda\leq \lambda'$. Moreover, $\lambda=\lambda'$ if and only if
    \begin{equation}\label{pre0}
        (PC'Q)_{ij}=(PCQ)_{ij}\qquad \hbox{for~}1\leq i, j\leq n \hbox{~with~} v_i\ne 0 \hbox{~and~} u_j\ne 0.
    \end{equation}
\end{thm}

\begin{proof}
    Multiplying the nonnegative vector $u$ in assumption~\ref{pre_thm_em2}

    to the right of both terms of~\ref{pre_thm_em1},
    \begin{equation}\label{pre1}
       PCQu\leq PC'Qu=\lambda'PQu,
    \end{equation}
 where the above equality follows by $Qu$ being eigenvector of $C'$ for $\lambda'$.
    Multiplying the nonnegative vector $v^T$ of $C$ in assumption
     ~\ref{pre_thm_em3} to the left of all terms  in (\ref{pre1}), we have
    \begin{equation}\label{pre2}
        \lambda v^TPQu=v^TPCQu\leq v^TPC'Qu=\lambda' v^TPQu,
    \end{equation}
    where the above first equality follows by $v^TP$ being
    left eigenvector of $C$ for $\lambda$.
    Now delete the positive term $v^TPQu$ by assumption \ref{pre_thm_em4} to obtain
        $\lambda\leq \lambda'$ and finish the proof of the first part.
        Assume that $\lambda=\lambda'$, so the inequality in (\ref{pre2}) is an equality.
        Especially $(PCQu)_i=(PC'Qu)_i$ for any $i$ with $v_i\not=0.$ Hence,
        $(PCQ)_{ij}=(PC'Q)_{ij}$ for any $i$ with $v_i\not=0$ and any $j$ with
        $u_j\not=0.$ Conversely, (\ref{pre0}) implies
        \begin{align*} &\lambda v^TPCQu=v^TPCQu=\sum_{i,j} v_i(PCQ)_{ij}u_j\\ =&
         \sum_{i,j} v_i(PC'Q)_{ij}u_j=v^TPC'Qu=\lambda'v^TPC'Qu,
         \end{align*} so $\lambda=\lambda'$ by (\ref{pre2}).
\end{proof}

@@

We will apply Theorem~\ref{pre_thm} by using two particular square matrices $P$ and $Q$ to obtain our main result. 
We use $[n-1]$ as notation of the set of integers from $1$ to $n-1$, which is $\{1,2,\cdots,n-1\}$.
Throughout this thesis we fix $k\in [n-1]$, let $E_{kn}$ denote the $n\times n$ binary matrix with a unique $1$
appearing in the position $k,n$ of $E_{kn}$. Now we apply the previous Theorem \ref{pre_thm} with $P=I$ and

\begin{equation} \label{Q_1}
Q=I+E_{kn}=\begin{pmatrix}
1 &   &        &       &        &   & 0 \\
  & 1 &        &       &        &   &   \\
  &   & \ddots &       &        &   &   \\
  &   &        & 1     &        &   & 1 \\
  &   &        &       &\ddots  &   &  \\
  &   &        &       &        &1  & 0 \\
0 &   &        &       &        &0  & 1
\end{pmatrix},
\end{equation}
so the matrix $PC'Q$ in assumption~\ref{pre_thm_em1} of Theorem~\ref{pre_thm} is
\begin{equation}\label{PC'Q}
  PC'Q=\begin{pmatrix}
         c'_{11} & c'_{12} & \cdots &  c'_{1k}+c'_{1n} \\
         c'_{21} & c'_{22} & \cdots &  c'_{2k}+c'_{2n} \\
         \vdots & \vdots & \ddots &  \vdots \\
         c'_{n1} & c'_{n2} & \cdots &  c'_{nk}+c'_{nn}
       \end{pmatrix},
\end{equation}
where $c'_{ij}$ denotes the $(i,j)$-entry of $C'$.

\begin{defn}%[$k$-rooted vector]
A column vector $v'=(v'_1,v'_2,\ldots,v'_n)^T$ is called {\it $k$-rooted} if $v'_{j} \geq 0$ for $1 \leq  j \leq n$ and $v'_k\geq v'_n$.
\end{defn}

The following Lemma is immediate from the above definition.

\begin{lem}\label{lem:rt_vec}
If $u=(u_1, u_2, \ldots, u_n)^T$, then
\begin{enumerate}[label=(\roman*)]
\item \label{lem:rt_vec:en1} $Qu$ is $k$-rooted  if and only if $u$ is nonnegative;
\item $u_k>0$ if and only if $(Qu)_k>(Qu)_n$.
\end{enumerate}
% \qed
\end{lem}

\begin{proof}
(i), (ii) follow immediate from the definition of $k$-rooted and
 $Qu=(u_1,\ldots, u_{k-1},u_k+u_n, u_{k+1}, \ldots,  u_n)^T$.
\end{proof}

Below is our first result, in which the first condition implies the first $n-1$ columns of $C'$ major
 the same columns of $C$, and the sum of $k$-th and $n$-th columns of $C'$ also majors that of $C$. The second and the third condition
 suggest that $C$ and $C'$ have nonnegative $k$-rooted eigenvectors. And the forth condition is simpler
  but with the same meaning in Theorem~\ref{pre_thm}

We need a notation of submatrix, which is taken from some columns and some rows of a matrix.

\begin{defn}
    For a matrix $C=(c_{ij})$ and subsets $\alpha$, $\beta$ of row indices and column
    indices of $C$, respectively, we use $C[\alpha|\beta]$ to denote the
    submatrix of $C$ with size $ |\alpha| \times |\beta| $ that has entries $c_{ij}$ for $i\in \alpha$
    and $j\in\beta$. We use the notation $C[i|j]$ for short of $C[[i]|[j]].$
\end{defn}

\begin{thm}\label{thm_main}
    Let $C=(c_{ij})$, $C'=(c'_{ij})$ be  $n\times n$ real matrices with real eigenvalues $\lambda$ and $\lambda'$ respectively.
Assume that
\begin{enumerate}[label=(\roman*)]
\item \label{thm_main:condition_i} $C[n|n-1]\leq C'[n|n-1]$ and $c_{ik}+c_{in}\leq c'_{ik}+c'_{in}$ for all $1\leq i\leq n$;
\item \label{thm_main:condition_ii} there exists a $k$-rooted eigenvector vector $v'=(v'_1, v'_2, \ldots, v'_n)^T$ of $C'$ for $\lambda'$;
\item \label{thm_main:condition_iii}there exists a nonnegative eigenvector vector $v^T=(v_1, v_2, \ldots, v_n)$ of $C$ for $\lambda$;
 \item \label{thm_main:condition_iv}$v^Tv'>0.$
\end{enumerate}
 Then $\lambda\leq \lambda'$.
Moreover, $\lambda=\lambda'$
if and only if
\begin{enumerate}[label=(\alph*)]
    \item \label{thm_main:equ_cond_a} $c_{ik}+c_{in}=c'_{ik}+c'_{in} \qquad$  for $1\leq i\leq n$ with $v_i>0$ and $v'_n>0;$
    \item \label{thm_main:equ_cond_b} $c'_{ij}=c_{ij}\qquad $for $1\leq i\leq n,~1\leq j\leq n-1,$ $j \neq k$ with $v_i\ne 0 $ and $v'_j>0$;
    \item \label{thm_main:equ_cond_c} $c'_{ik}=c_{ik} \qquad $  for $1\leq i \leq n$ with  $v_i>0$ and $ v'_{k}>v'_n$
\end{enumerate} 
\end{thm}

@@

We are interested in the matrices $C'$ that have $k$-rooted eigenvectors.
Motivated by the condition (i) of Theorem 2.3, we provide the following two definitions.
The first is the definition of $(k,n)$-sum.
\begin{defn}
    For an $n \times n$ matrix $C'=(c'_{ij})$, the $(k, n)$-{\it sum vector} of $C'$ is the vector
     obtained from the sum of the $k$-th and  $n$-th columns of $C'$.
\end{defn}

Note that the last column of $C'Q$ is the $(k, n)$-sum vector of $C'$.
Below is the definition of $k$-rooted matrix.
\begin{defn}\label{m_rooted}
    A  matrix $C'=(c'_{ij})$ is called {\it $k$-rooted}  if for
     each $i\not=k, n$ the  $i$-th column of $C'$  is $k$-rooted
     and the $(k,n)$-sum vector of $C'$ is $k$-rooted.
\end{defn}

We need a simple lemma for later use.

\begin{lem}
$Q^{-1}=I-E_{kn}.$
\end{lem}

\begin{proof}
Since $$Q(I-E_{kn}=(I+E_{kn})(I-E_{kn})=I-E_{kn}+E_{kn}-E_{kn}E_{kn}=I,$$ we have
$Q^{-1}=I-E_{kn}$.
\end{proof}

The matrix $Q^{-1}$ is explicitly written as
    $$Q^{-1}=I-E_{kn}=
\begin{pmatrix}
1 &   &        &       &        &   & 0 \\
  & 1 &        &       &        &   &   \\
  &   & \ddots &       &        &   &   \\
  &   &        & 1     &        &   & -1 \\
  &   &        &       &\ddots  &   &  \\
  &   &        &       &        &1  & 0 \\
0 &   &        &       &        &0  & 1
\end{pmatrix},$$ 
and if $C'=(c'_{ij})$ then $Q^{-1}C'Q$ has the form 
\begin{equation}\label{e6}
\begin{pmatrix}
            c'_{11}     & c'_{12} & \cdots     & c'_{1\ n-1} & c'_{1k}+c'_{1n} \\
            \vdots      & \vdots  & \vdots     & \vdots      & \vdots\\
            c'_{(k-1) 1}   & c'_{(k-1)  2}           & \cdots     & c'_{(k-1) (n-1)} & c'_{(k-1) k}+c'_{(k-1) n} \\
            c'_{k1}-c'_{n1} & c'_{k2}-c'_{n2} &\cdots      &c'_{k (n-1)}-c'_{n (k-1)}& c'_{kk}+c'_{kn}-c'_{nk}-x'_{nn}\\
            c'_{(k+1) 1}   & c'_{(k+1) 2}           & \cdots     & c'_{(k+1) (n-1)} & c'_{(k+1) k}+c'_{(k+1)
             n} \\
            \vdots              & \vdots & \ddots              & \vdots & \vdots \\
            c'_{n1}     & c'_{n2} & \cdots             & c'_{n (n-1)} & c'_{nk}+c'_{nn} \\
        \end{pmatrix}.\end{equation}

The following lemma shows that a $k$-rooted matrix has a $k$-rooted eigenvector.

\begin{lem}\label{lma_m_rooted}
    Let $C'=(c'_{ij})$ be an $n\times n$ nonnegative matrix. Then the following (i)-(ii) hold.
        \begin{enumerate}[label=(\roman*)]
            \item \label{lma_m_rooted_cond1} $C'$ is a $k$-rooted matrix if and only if $Q^{-1}(C'+dI)Q$ is nonnegative for some $d\geq 0$, where $I$ is the $n\times n$ identity matrix. 
            \item \label{lma_m_rooted_cond2} If $C'$ is $k$-rooted then there exists a  $k$-rooted eigenvector $v'$ of $C'$  for $\rho(C')$.
        \end{enumerate}
\end{lem}

\begin{proof}
\begin{enumerate}
  \item[(i)] The matrix $Q^{-1}C'Q$ has $ij$ entry
$$(Q^{-1}C'Q)_{ij}=\left\{
                     \begin{array}{ll}
                       c'_{ij}, & \hbox{if $i\not=k$ and $j\not=n$;} \\
                       c'_{kj}-c'_{nj}, & \hbox{if $i=k$ and $j\not=n$;} \\
                       c'_{ik}+c'_{in}, & \hbox{if $i\not=k$ and $j=n$;} \\ 
                       c'_{kk}+c'_{kn}-c'_{nk}-c'_{nn}, & \hbox{if $i=k$ and $j=n$},
                     \end{array}
                   \right.$$
as shown in (\ref{e6}). 
Hence $Q^{-1}(C'+dI)Q$ is nonnegative if and only if $C'$ is $k$-rooted by the definition of nonnegative matrix and $k$-rooted matrix.
  \item[(ii)] Suppose $C'$ is $k$-rooted. Choose $d\geq 0$ such that $Q^{-1}(C'+dI)Q$ is nonnegative.  
Let $u$ be a nonnegative eigenvector of $Q^{-1}(C'+dI)Q=Q^{-1}C'Q+I$ for $\rho(C'+dI)=\rho(C')+d.$
Note that $Q^{-1}C'Qu=\rho(C')u,$ and $Qu$ is $k$-rooted by Lemma~\ref{lem:rt_vec}. Hence $v'=Qu$ is what we want. 
\end{enumerate}
\end{proof}

Note that Theorem~\ref{thm_main} depends on eigenvectors. The following is an eigenvector-free Theorem.

\begin{thm}\label{thm:conclusion}
    Let $C$ be an $n\times n$ nonnegative irreducible matrix and $C'$ be an $n\times n$ $k$-rooted matrix such that $C'Q$ majorizes $CQ$. 
 Then $\rho(C)\leq \rho(C')$. 
\end{thm}
